Mining Scenario/Transaction Pairs HW/SW Co-verification Using Symbolic Execution

### 2017-04-01 paper.tex
updated compiling with ACM class
intro

\documentclass[format=acmsmall, review=false, screen=true]{acmart}

@@

\begin{document}

\title{Mining Scenario/Transaction Pairs HW/SW Co-verification Using Symbolic Execution}

@@

%
%===============================================================================
\begin{abstract}
%===============================================================================
%
Time-to-market pressure requires early development of firmware and other
drivers for circuitry well before RTL is finished.  A key risk is divergence
of the hardware RTL from the behavior expected by the software.  Formal
methods can deliver proof that the RTL and the firmware match, but are
difficult to scale up.  We propose to identify partitions of the firmware and
the RTL that follow transaction boundaries by means of path-based symbolic
simulation with custom path-pruning.  We have implemented this approach in
our tool \verifox.  We have experimentally compared \verifox with a
conventional approach, which constructs a single monolithic SAT/SMT
instance, and observe speedups of an order of magnitude over conventional 
approach for proving safety properties as well as detecting firmware/hardware bugs.
%
\end{abstract}

%===============================================================================
\section{Introduction}
%===============================================================================
Hardware/software co-verification is broadly performed at two different
phases in the design cycle: pre-RTL and post-RTL.  In the pre-RTL phase,
engineers write high-level software (SW) models for Intellectual Property (IP), 
which are commonly known as {\em Transaction Level Model}s
(TLM)~\cite{codes14}.  TLMs are typically implemented using the SystemC TLM
library~\cite{tlm} or as plain C programs.  The goal of a TLM is to capture
enough functionality of the hardware (HW) such that it is possible to execute the
SW (drivers/firmware) against it.  Consequently, the TLMs are
incomplete, as they are designed to contain only those behaviors that are
observable by the SW (the drivers or firmware).  In particular, TLMs
are often not cycle-accurate.  These models are used to verify high-level
functionality of the system well before the RTL is
available~\cite{codes14,codes15}. Co-verification of the TLM and the
SW is typically performed by means of testing.

@@

Note that whenever we refer to ``hardware'' from this point onwards, we mean
an RTL implementation and not a TLM.  A firmware services the requests from
the application software layer or its interacting hardware components. 
Similarly, the hardware services the requests from the firmware or the
physical environment.  Thus, the firmware and its interacting hardware
components can be seen as a set of services.  In this paper, we call a
hardware RTL service a {\em Hardware Transaction} $(\mathcal{HT})$.  Closest
to our notion of $\mathcal{HT}$ is the micro-architecture-level transaction
by Mahajan et al.~\cite{mcbmq}.  We call a FW service a {\em Firmware
Transaction} ($\mathcal{FT}$).  $\mathcal{FT}$ is an untimed finite-state
machine that is manifested as control-flow paths in the FW, similar to
transactions in TLM~\cite{codes14}.  A~{\em scenario} $(\mathcal{S})$ in a
co-verification environment typically configures the hardware in a way that
restricts the interaction between $\mathcal{FT}$ and $\mathcal{HT}$.  In
other words, given a $\mathcal{S}$, the firmware only exercises a fragment
of the HW state-space.  This renders many interactions between
$\mathcal{HT}$ and $\mathcal{FT}$ infeasible.  A {\em Scenario/Transaction}
pair is a set of valid interactions between $\mathcal{HT}$ and
$\mathcal{FT}$ for a particular~$\mathcal{S}$.  Thus, it is necessary to
automatically mine Scenario/Transaction pairs to scale HW/SW co-verification
to SoC level designs.  A conventional, monolithic approach, e.g., an
approach based on BMC~\cite{CKY03}, cannot exploit this infeasibility.  The
key idea in this paper is to use path-based symbolic execution to
automatically infer such Scenario/Transaction pairs and check their validity
using backend solvers.

@@

%===============================================================================
%\subsection{Contributions}
%===============================================================================
In this paper, we make the following contributions:
\begin{enumerate}
\item {\em Co-verification using Symbolic Execution:} 
In this paper, we introduce a notion of {\em transaction} 
in HW RTL and FW models. We present a unified FW-HW model, 
called {\em Program Netlist}, in C, for the co-verification task. 
To this end, we present a path-based symbolic execution algorithm that is 
able to automatically find scenario/transaction pairs from a 
program netlist and check their validity. 
To the best of our knowledge, this is the first attempt to 
perform path-based symbolic execution of 
C (FW) and Verilog RTL (HW) models by synthesizing 
them to a unified program netlist model.

>>>>

\documentclass[sigconf]{acmart}

@@

\title{Mining Scenario/Transaction Pairs HW/SW Co-verification Using Symbolic Execution}
%===============================================================================
\begin{abstract}
%===============================================================================
%
Time-to-market pressure requires early development of firmware and other
drivers for circuitry well before RTL is finished.  A key risk is divergence
of the hardware RTL from the behavior expected by the software.  Formal
methods can deliver proof that the RTL and the firmware match, but are
difficult to scale up.  We propose to identify partitions of the firmware and
the RTL that follow transaction boundaries by means of path-based symbolic
simulation with custom path-pruning.  We have implemented this approach in
our tool \verifox.  We have experimentally compared \verifox with a
conventional approach, which constructs a single monolithic SAT/SMT
instance, and observe speedups of an order of magnitude over conventional 
approach for proving safety properties as well as detecting firmware/hardware bugs.
%
\end{abstract}

\begin{document}
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\pagestyle{plain} % removes running headers
\setcopyright{none}

@@

%
%===============================================================================
\section{Introduction}
%===============================================================================
Hardware/software co-verification is broadly performed at two different
phases in the design cycle: pre-RTL and post-RTL.  In the pre-RTL phase,
engineers write high-level software (SW) models for Intellectual Property (IP), 
which are commonly known as {\em Transaction Level Model}s
(TLM)~\cite{codes14}.  TLMs are typically implemented using the SystemC TLM
library~\cite{tlm} or as plain C programs.  The goal of a TLM is to capture
enough functionality of the hardware (HW) such that it is possible to execute the
SW (drivers/firmware) against it.  Consequently, the TLMs are
incomplete, as they are designed to contain only those behaviors that are
observable by the SW (the drivers or firmware).  In particular, TLMs
are often not cycle-accurate.  These models are used to verify high-level
functionality of the system well before the RTL is
available~\cite{codes14,codes15}. Co-verification of the TLM and the
SW is typically performed by means of testing.

@@

A firmware only exercise some fragment of hardware state-space in a typical hardware/software co-design~\cite{dac17,taa,ibm-other}.  However, there is no automatic hardware/software co-verification tool that can exploit this interaction behavior.  For building effective and scalable formal co-verification tool, it is therefore imperative to automatically mine only relevant scenario/transaction pairs that occurs in practical co-designs.  In this paper, we attempt to bridge this gap by presenting a technique that automatically mine relevant HW/SW interaction behavior using symbolic execution.
A firmware services the requests from
the application software layer or its interacting hardware~\footnotesize{Note that whenever we refer to ``hardware'' from this point onwards, we mean
an RTL implementation and not a TLM.}  components. 
Similarly, the hardware services the requests from the firmware or the
physical environment.  Thus, the firmware and its interacting hardware
components can be seen as a set of services.  In this paper, we call a
hardware RTL service a {\em Hardware Transaction} $(\mathcal{HT})$.  Closest
to our notion of $\mathcal{HT}$ is the micro-architecture-level transaction
by Mahajan et al.~\cite{mcbmq}.  We call a FW service a {\em Firmware
Transaction} ($\mathcal{FT}$).  $\mathcal{FT}$ is an untimed finite-state
machine that is manifested as control-flow paths in the FW, similar to
transactions in TLM~\cite{codes14}.  A~{\em scenario} $(\mathcal{S})$ in a
co-design typically configures the co-verification environment in a way that
restricts the interaction between $\mathcal{FT}$ and $\mathcal{HT}$.  In
other words, given a scenario $\mathcal{S}$, the FW only exercises a fragment
of the HW state-space.  This renders many interactions between
$\mathcal{HT}$ and $\mathcal{FT}$ infeasible. For example, a scenario in \emph{Universal Asynchronous Receiver Transmitter} (UART) IP can configure the UART in \emph{transmitter} mode which restricts the communication between the FW and HW to set of services that only transfers data from the FW to the serial output port of the transmitter HW while completely bypassing the receiver HW.    A {\em Scenario/Transaction}
pair is a set of valid interactions between $\mathcal{HT}$ and
$\mathcal{FT}$ for a particular scenario~$\mathcal{S}$.  
It is necessary to automatically mine Scenario/Transaction pairs for scalable and effective co-verification.

A conventional, monolithic approach, e.g., an
approach based on BMC~\cite{CKY03}, cannot exploit this infeasibility.  The
key idea in this paper is to use path-based symbolic execution to
automatically infer such Scenario/Transaction pairs and check their validity
using backend solvers.

@@

%===============================================================================
%\subsection{Contributions}
%===============================================================================
In this paper, we make the following contributions:
\begin{enumerate}
\item Translating Hardware/Software Co-designs to an unified 
sequential software model, represented in C.  
verification using Software Model Checkers.
\item {\em Co-verification using Symbolic Execution:} 
In this paper, we introduce a notion of {\em transaction} 
in HW RTL and FW models. We present a unified FW-HW model, 
called {\em Program Netlist}, in C, for the co-verification task. 
To this end, we present a path-based symbolic execution algorithm that is 
able to automatically find scenario/transaction pairs from a 
program netlist and check their validity. 
To the best of our knowledge, this is the first attempt to 
perform path-based symbolic execution of 
C (FW) and Verilog RTL (HW) models by synthesizing 
them to a unified program netlist model.
